---
title: "CyTOF Data Analysis"
output:
  github_document: default
  html_document:
    df_print: paged
---
### The example data sets
We use a of CyTOF data from Environmental Science:Nano Paper, My et al.2020. There are 7 samples of PBMC from human. Treated samples are coated with Ag nano pariticle in various method.We use a leukocyte sample which are manually gated and concatenated in FlowJo Software.  
The original data is available from the link below.  
https://drive.google.com/drive/u/0/folders/1hD8bfT8z-9y8lGRDH0sTb_Jn12xw0S4X


### Data preprocessing 

##### Read fcs file
The FCS format is a binary data file standard originally developed for storage of flow cytometry data. As well as the ion count matrix, it has slots for experiment information, channel information, etc. 
It is recommended that you set the current path using setwd() in console and put the file in that directory.
```{r, warning=F}
library(flowCore)
fcs <- read.FCS(filename="../data/Leukocytes.fcs", transformation=F, truncate_max_range = F)

```

##### Extract expression matrix
```{r}
exprs <- fcs@exprs
head(exprs)
```

##### Explore the parameters
fcs file also contains metadata, such as the expression range of each channel.
```{r}
fcs@parameters@data
```

Make name of columns human readable using information in the parameter data slot
```{r}
markers <- gsub(pattern = "_$", replacement = "", x = as.vector(fcs@parameters@data$desc))
markers <- gsub(pattern = ".*_", replacement = "", x = markers)
markers <- gsub(pattern = "-", replacement = ".", x = markers)
colnames(exprs) <- markers
surface_markers <- c("CD3", "CD4", "CD8a", "CD11a", "CD16", "CD44", "CD45RA", "CD45RO", "HLA.DR", "CCR7")
```

##### Data transformation
Mass cytometry data is usuallly ArcSinh transformed. This transformation will retain linearity in the low end of the count spectrum, but it resembles a log transformation in the high end.  Transformations are done primarily for visualization purposes
```{r}
cofactor <- 5
exprs[, surface_markers] <- asinh(exprs[, surface_markers] / cofactor)
```

### Dimensionality Reduction

##### PCA
For mass cytometry data, a PCA will often capture 40-50% of the variance on the first 2-3 PCs, but since the distance between events is linear, data existing on a nonlinear manifold will be misrepresented (Chester & Maecker  et al., 2015). 
```{r}
pca <- prcomp(exprs[, surface_markers], scale. = TRUE)
library(ggplot2)
ggplot(as.data.frame(pca$x), aes(x = PC1, y = PC2)) + 
  geom_point(color = "grey", alpha = 0.6, size = 0.5) + 
  theme_bw() 
```

##### t-SNE
The t-Distributed Stochastic Neighbor Embedding (t-SNE) dimensionality reduction algorithm is a popular tool for dimensionality reduction of cytometry data (van der Maaten and Hinton et al., 2008).  
Before running t-SNE, it is necessary to remove identical events (if any) from the sample (duplicate rows).We set a seed for get reproducible results.
```{r}
library(Rtsne)
table(duplicated(exprs[, surface_markers]))
exprs <- exprs[!duplicated(exprs[, surface_markers]), ]
set.seed(42)
tsne <- Rtsne(exprs[, surface_markers])
ggplot(as.data.frame(tsne$Y), aes(x = V1, y = V2)) +
  geom_point(color = "grey", alpha = 0.6, size = 0.5) +
  xlab('tSNE1') + ylab('tSNE2') +
  theme_bw()
```

##### UMAP
UMAP works in a dimensionality reduction method that is more similar to t-SNE than PCA. However, it has several advantages over t-SNE. The first one is speed. For a small dataset there may not be a significant difference, bu you will surely notice it for a million cells. Furthermore, UMAP preserves global distances in the data a lot better than t-SNE.  We set a seed for get reproducible results.
```{r, message=F}
library(uwot)
umap <- umap(exprs[, surface_markers], min_dist = 0.2)
ggplot(as.data.frame(umap), aes(x = V1, y = V2)) +
  geom_point(color = "grey", alpha = 0.6, size = 0.5) +
  xlab('UMAP1') + ylab('UMAP2') +
  theme_bw()
```

##### Visualize tSNE plot with several marker intensities
```{r, message=F}
library(reshape2) 
library(cowplot) 
obj <- cbind.data.frame(exprs, tsne$Y) 
colnames(obj)[(ncol(obj)-1):ncol(obj)] <- c('tSNE1', 'tSNE2') 
colnames(obj) <- make.unique(names(obj))
p <- list() 
for (c in c("CD4", "CD8a", "CD45RA", "CCR7")) { 
  p[[c]] <- ggplot(obj, aes(x = tSNE1, y = tSNE2)) + 
    geom_point(aes_string(color = c), size = 0.5) + 
    theme_bw() 
} 
plot_grid(plotlist = p, ncol = 2) 
```

### Clustering 
##### Phenograph
Phenograph is an algorithm which directly assigns single cells to a particular cluster while taking the all measured dimensions into account. Phenograph employs a k-nearest neighbor graph (k-NNG), in which each cell is a node that is connected to its nearest neighbors by edges, to cluster cells into phenotypically similar groups - cellular populations. Phenograph also requires removal of duplicates.   
Phenograph is implemented in the R package “cytofkit”.
```{r,warning=F, message=F}
library(cytofkit) 
clusters_pg <- cytof_cluster(xdata = exprs[,surface_markers], method = "Rphenograph") 
```

It is a vector that contains the number of clusters of entire cell.  
```{r}
head(clusters_pg)
```


##### Visualize phenograph results on tSNE
```{r}
library(RColorBrewer)
getPalette = colorRampPalette(brewer.pal(9, "Set1")) 
cols <- getPalette(max(clusters_pg)) 
plot_df <- cbind.data.frame(obj, clusters_pg) 
plot_df$clusters_pg <-as.factor(plot_df$clusters_pg) 
ggplot(plot_df, aes(x = tSNE1, y = tSNE2)) + 
  geom_point(aes(color = clusters_pg), alpha = 0.6, size = 0.5) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2), 
                               ncol = 2, title = 'Cluster')) + 
  scale_color_manual(values = cols) + 
  theme_bw() 

```

### References
- My Kieu Ha (2020). Mass cytometric study on the heterogeneity in cellular association and cytotoxicity of silver nanoparticles in primary human immune cells. Environmental Science:Nano, 2020,7, 1102-1114. doi:10.1039/C9EN01104H
- Data scientist’s primer to analysis of mass cytometry data, The Cancer Omics Group, https://biosurf.org/cytof_data_scientist.html#1_preamble


