---
title: "Importing gating informations and assigning cell type to clusters"
output:
  github_document: default
  html_document:
    df_print: paged
---
# Import gating information



### Import gating information from cytobank
With gatingML file and related fcs files, you can get the gatingset object with CytoML
```{r, message=F, warning=F}
library(flowWorkspace)
library(CytoML)
library(flowCore)

#full path of gatingML and fcsfile 
xmlfile <- "../data/CytExp_337288_Gates_v6.xml"
fcsfile <- "../data/2019 1 29  Control_Leukocyte.fcs"
#Get gatingset object from gatingML file and fcs file.
gs <- cytobank_to_gatingset(xmlfile, fcsfile)
#Get gated flow data from gatingset
fs <-  gs_pop_get_data(gs)
```

Also, you can get the populations of gated cell type and gating hierarchy
```{r, message=F, warning=F}
# Extract the population statistics
gs_pop_get_count_fast(gs, statType = "count")
```

```{r, message=F, warning=F}
plot(gs)
```

```{r,echo=FALSE}
rm(list=ls())
```


### Import Gating information from cytobank
With Flowjo workspace file and related fcs files, you can get the gatingset object with CytoML. Both files must be in the same directory. 
```{r, message=F, warning=F}
#Import Flowjo workssace
ws <- open_flowjo_xml("../data/concat_leukocytes.wsp")
gs <- flowjo_to_gatingset(ws, name=1)
#Get gated flow data from gatingset
fs <- gs_pop_get_data(gs)
```

When importing gating information, CytoML transforms several channels in flow data which are used in gating as they are written in gating information file. To treat all channels equally, we have to transform flow data inversely and make it raw data. 

```{r, message=F, warning=F}
# Inverse transformation
trans.inv <- gh_get_transformations(gs[[1]], inverse = TRUE) 
translist <- transformList(names(trans.inv), trans.inv)
fs.raw <- transform(fs, translist)
# Extract expression matrix from flow data 
exprs <- exprs(fs.raw[[1]])
exprs <- as.data.frame(exprs)
```


##### Change marker name easy to read 
```{r, message=F, warning=F}
(params <- parameters(fs.raw[[1]])@data)
```

Depending on the fcs file, the name of the channel can vary. It would be good process to change channel names by referring to description of parameter and make it easy to read.

```{r, message=F, warning=F}
#Convert Marker names by referring to description of parameter  
for(i in colnames(exprs)){
  
  if(i %in% params[,1]){
    colnames(exprs)[colnames(exprs)==i] <- params[params$name==i,2]
  }
}
#Remove channles without description 
exprs <- exprs[,!is.na(colnames(exprs))] 
#Leave the name of target antibody only
colnames(exprs) <- gsub(pattern = "^[0-9]*", replacement ="",x=colnames(exprs))
colnames(exprs) <- gsub(pattern = "-", replacement =".",x=colnames(exprs))
colnames(exprs) <- gsub(pattern = "_$", replacement ="",x=colnames(exprs))
colnames(exprs) <- gsub(pattern = ".*_", replacement ="",x=colnames(exprs))
colnames(exprs) <- make.unique(colnames(exprs),sep = "_")
```

##### Annotating sample condtion and manually gated cell type 
In this sample, sample condition nodes have depth 1 and cell type nodes are leaf node at the tree of gating hierarchy.  


To annotate these conditions, we use ‘gh_pop_get_indices’ function that returns a logical vector that describes each event in a sample is included by certain gate.  
```{r}
#Plot gating hierarchy
plot(gs,fontsize=30)
```

We can annotate sample condition by selecting nodes which have depth 1 and using ‘gh_pop_get_indices’ function.
```{r, message=F, warning=F}
#Annotate Sample names
exprs["sample"] <- rep("NA",nrow(exprs))
path <- gs_get_pop_paths(gs)
for(i in 1:length(path)){
  split_path <- strsplit(path[i],"/")
  if(length(split_path[[1]])==2){
    exprs[gh_pop_get_indices(gs,path[i]),"sample"] <- split_path[[1]][2]
  }
}
```

We can annotate cell type by selecting leaf nodes without child node and using ‘gh_pop_get_indices’ function. The cells that are not included in leaf nodes are annotated as ‘ungated’
```{r, message=F, warning=F}
#Annotate cell types
exprs["celltype"] <- rep("ungated",nrow(exprs))
exprs["node"] <- rep("ungated",nrow(exprs))
for(i in gs_get_leaf_nodes(gs)){
  #i <- gsub(".*/", "", i)
  cellType <- gsub(".*/", "", i)
  exprs[gh_pop_get_indices(gs,i),"celltype"] <- cellType
  exprs[gh_pop_get_indices(gs,i),"node"] <- i
}

```

```{r, message=F, warning=F}
table(exprs[,"sample"],exprs[,"celltype"])
```
To treat easily, conver these character values to factor values

```{r, message=F, warning=F}
exprs[,c("celltype","node","sample")] <- lapply(exprs[,c("celltype","node","sample")], as.factor)
str(exprs)
```

##### Benefits of CytoML

- identifying gating information cell by cell in R
  - For example, It can be used for learning of the automatic gating and calculation of gating populations.

- Free transition of working environment between different platforms. 




# Assign cluster to celltype by calculating F1 score matrix

Select the markers used for finding population in manual gating as markers for clustering.
```{r, message=F, warning=F}
interest_markers <- c("NK1.1", "Gr.1", "CD11c", "Mac.1", "CD3e", "TCRb", "CD19", "B220")
interest_markers %in% colnames(exprs)
```

Cluster the cells with Phenograph method. 
```{r, message=F, warning=F}
library(cytofkit)
clusters_pg <- cytof_cluster(xdata = exprs[,interest_markers], method = "Rphenograph") 
```
### Measure F1 Score matrix
Using manually gated cell type as actual value and cluster number as predicted value, calculate precsion, recall, F1 score.Then we can make a F1 Score matrix 
```{r, message=F, warning=F}

tbl_celltype <- table(exprs[,"celltype"])
tbl_cluster <- table(clusters_pg)
F1_mat <- matrix(NA, nrow=length(tbl_cluster),ncol=length(tbl_celltype))

for (i in 1:length(tbl_cluster)){
  for (j in 1:length(tbl_celltype)){
    
    true_positives <- sum(clusters_pg == i &
                          exprs[,"celltype"] == levels(exprs[,"celltype"])[j])
    detected <- sum(clusters_pg == i)
    truth <- sum(exprs[,"celltype"] == levels(exprs[,"celltype"])[j])
    
    #calculate precision, recall and F1 Score
    precision_ij <- true_positives / detected
    recall_ij <- true_positives / truth
    F1_ij <- 2 * (precision_ij * recall_ij) / (precision_ij + recall_ij)
    
    if (F1_ij == "NaN") F1_ij <- 0
    
    F1_mat[i, j] <- F1_ij
  }
}
rownames(F1_mat) <- names(tbl_cluster)
colnames(F1_mat) <- names(tbl_celltype)
F1_mat
```
### Assignment method 1: 1-1 mapping by Hungarian algorithm
##### Hungarian algorithm
The Hungarian matching algorithm, also called the Kuhn-Munkres algorithm, is a O(n^3) algorithm that can be used to find maximum-weight matchings in bipartite graphs, which is sometimes called the assignment problem. Hungarian algorithm is implemented by using solve_LSAP function. Using this algorithm, we can match the optimal cluster to cell type matching that sum of the F1 score maximize.
```{r, message=F, warning=F}
# use transpose matrix (Hungarian algorithm assumes n_rows <= n_cols)
F1_mat_trans <- t(F1_mat)
if (nrow(F1_mat_trans) <= ncol(F1_mat_trans)) {
  # if fewer (or equal no.) true populations than detected clusters, can match all true populations
  labels_matched <- clue::solve_LSAP(F1_mat_trans, maximum = TRUE)
  # use row and column names since some labels may have been removed due to unassigned cells
  labels_matched <- as.numeric(colnames(F1_mat_trans)[as.numeric(labels_matched)])
  names(labels_matched) <- rownames(F1_mat_trans)
  
} else {
  # if fewer detected clusters than true populations, use transpose matrix and assign
  # NAs for true populations without any matching clusters
  labels_matched_flipped <- clue::solve_LSAP(F1_mat, maximum = TRUE)
  # use row and column names since some labels may have been removed due to unassigned cells
  labels_matched_flipped <- as.numeric(rownames(F1_mat_trans)[as.numeric(labels_matched_flipped)])
  names(labels_matched_flipped) <- rownames(F1_mat)
  labels_matched <- rep(NA, ncol(F1_mat))
  names(labels_matched) <- rownames(F1_mat_trans)
  labels_matched[as.character(labels_matched_flipped)] <- as.numeric(names(labels_matched_flipped))
}
labels_matched
```

Since it is 1-to-1 matching, there are many cluster that are not assigned.
```{r, message=F, warning=F}
pop <- rep(NA, nrow(exprs))
for (i in 1:length(pop)){
pop[i] <- ifelse(clusters_pg[i] %in% labels_matched, names(which(labels_matched==clusters_pg[i])), "unassigned")
}
exprs[,"celltype_1to1"] <- pop
table(exprs[,"celltype_1to1"])
```


### Assignment method 2: n-1 mapping by finding maximum F1 Score
Also using F1 score matrix, you can easily assign cell type to cluster by matching cell type with the largest F1 score for each cluster.   
```{r, message=F, warning=F}
labels_matched <- rep(NA, nrow(F1_mat))
for(i in 1:nrow(F1_mat)){
  labels_matched[i] <- names(which.max(F1_mat[i,]))
}
```


```{r, message=F, warning=F}
exprs[,"celltype_nto1"] <- labels_matched[clusters_pg]
table(exprs[,"celltype_nto1"])
```


